A Fast Unified Model for Parsing and Sentence Understanding

### 2016-06-07 writing/hard_stack_paper/paper.tex
CM's changes

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces the required parse structure on the fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser and saving its additional computational cost. Secondly, it supports batched computation for both parsed and unparsed sentences, yielding dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid architecture for handling local linear context in sentence interpretation. This model is a basically plausible model of human sentence processing and yields substantial accuracy gains over pure sequence- or tree-based models. % TRIMMABLE

@@

The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. The parser is initialized with the stack empty and the buffer containing the tokens $\mathbf x$ of the sentence in order. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. It next proceeds through the transition sequence, where each transition $a_t$ selects one of the two following operations. Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\vspace*{-1ex} % dirty spacing trick that somehow seemed needed right now
%TRIMMABLE
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it on to the top of the stack.
\vspace*{-1ex}
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back on to the stack.
\end{description}

@@

The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
     {\sigma\vphantom{\vec{i}}}
     {\sigma\vphantom{\vec{f}_l}}
     {\sigma\vphantom{\vec{f}_r}}
     {\sigma\vphantom{\vec{o}}}
    {\text{tanh}\vphantom{\vec{g}}}
\left(
W_{\text{comp}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
+ \vec{b}_{\text{comp}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector-valued input argument which is either empty or comes from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack. Each vector-valued variable listed is of dimension $D$ except $\vec{e}$, of the independent dimension $D_{\text{tracking}}$.

@@

We use word representations based on the standard 300D vectors provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to map each input word vector $\vec{x}_{\text{GloVe}}$ into a vector pair $\langle \vec{h}, \vec{c}\rangle$ that is stored in the buffer:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher (unnormalized) probability. The prediction function is trained to mimic the decisions of an external parser. These decisions are used as the inputs to the model during training. For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus. We did not find scheduled sampling \citep{bengio2015scheduled}---having the model use its own transition decisions sometimes at training time---to help.

@@

\begin{table}[t]
\centering
\begin{tabular}{cc @{\hspace*{2em}}c @{\hspace*{2em}} l}
  \toprule
  $t$ & $a_t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} \\
  \midrule
  0 & \shift &  & \underline{\hphantom{0} \hphantom{0}} \\
  1 & \shift & \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} \\
  2 & \shift & \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} \\
  3 & \reduce & \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} \\
  4 & \reduce & (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} \\
  5 & & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} \\
  \bottomrule
\end{tabular}
\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column. $S$ is shown in the third column and represents the top of the stack at each step $t$. The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
\label{tbl:thin-stack}
\end{table}

@@

A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep in an example, any element of which may be involved in later computations. A na\"ive backpropagation implementation would then require storing each of the $O(N)$ stacks for a backward pass, leading to a per-example space requirement of $O(NTD)$ floats. This requirement is prohibitively large for significant batch sizes or sentence lengths $N$. Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the top of the stack.

@@

\begin{figure}
\centering
\resizebox{1\linewidth}{!}{
\begin{tikzpicture}
\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
             xticklabels={0,32,64,128,256,512,1024,2048}, xmode=log, log basis x=2,
             legend style={at={(0.55, 0.5)},anchor=north west}]
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.tsv};
  \addlegendentry{Thin-stack GPU}
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.tsv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.tsv};
  \addlegendentry{RNN}
\end{axis}
\end{tikzpicture}
}
\caption{Feedforward speed comparison.}
\label{fig:speed}
\end{figure}

Each model is restricted to run on sentences of 30 tokens or fewer. We fix the model dimension $D$ and the word embedding dimension at 300. We run the CPU performance test on a 2.20-GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled. We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.

@@

Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of some model to extract broadly informative representations of sentence meaning. In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.

SNLI is a corpus of 570k human-labeled pairs of scene descriptions like the one above. We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing. SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.

@@

\begin{table*}[t]\small
\begin{center}
\begin{tabular}{lrlrrrr}
\toprule
Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
\midrule 
Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
\bottomrule
\end{tabular}
\end{center}
\caption{
\label{tab:hyperparams}
Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log-uniform.
}
\end{table*}

@@

Table~\ref{tab:results} shows our results on SNLI inference classification. For the full SPINN, we also report a measure of agreement between this model's parses and the parses included with SNLI, calculated as classification accuracy over transitions averaged across timesteps.

@@

We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses. We suspect that it is acting somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.

@@

For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations \citep[as in][]{grefenstette2015learning,joulin2015inferring}. This would make it possible for the model to learn to parse using guidance from the semantic representation objective, essentially allowing it to learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.

@@

We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant nos.~BCS 1456077 and IIS 1514268. Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation. We also thank Kelvin Gu, Noah Goodman, and many others in the Stanford NLP group for helpful comments.

>>>>

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces the required parse structure on the fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser at nearly no additional computational cost. Secondly, it supports batched computation for both parsed and unparsed sentences, yielding dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid architecture for handling local linear context in sentence interpretation. This model is a basically plausible model of human sentence processing and yields substantial accuracy gains over pure sequence- or tree-based models. % TRIMMABLE

@@

The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. The parser is initialized with the stack empty and the buffer containing the tokens $\mathbf x$ of the sentence in order. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. It next proceeds through the transition sequence, where each transition $a_t$ selects one of the two following operations. Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\vspace*{-1ex} % dirty spacing trick that somehow seemed needed right now
%TRIMMABLE
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it on to the top of the stack.
\vspace*{-1ex}
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them, and pushes the result back on to the stack.
\end{description}

@@

The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of an active state representation $\vec{h}$ and a memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
     {\sigma\vphantom{\vec{i}}}
     {\sigma\vphantom{\vec{f}_l}}
     {\sigma\vphantom{\vec{f}_r}}
     {\sigma\vphantom{\vec{o}}}
    {\text{tanh}\vphantom{\vec{g}}}
\left(
W_{\text{comp}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
+ \vec{b}_{\text{comp}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
where $\sigma$ is the sigmoid activation function, $\odot$ is the elementwise product, the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$ are the two input tree nodes popped off the stack, and $\vec{e}$ is an optional vector-valued input argument which is either empty or comes from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). The result of this function, the pair $\langle\vec{h}, \vec{c}\rangle$, is placed back on the stack. Each vector-valued variable listed is of dimension $D$ except $\vec{e}$, of the independent dimension $D_{\text{tracking}}$.

@@

We use word representations based on the 300D vectors provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to map each input word vector $\vec{x}_{\text{GloVe}}$ into a vector pair $\langle \vec{h}, \vec{c}\rangle$ that is stored in the buffer:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

At test time, the model uses whichever transition (i.e., \shift\ or \reduce) is assigned a higher (unnormalized) probability. The prediction function is trained to mimic the decisions of an external parser. These decisions are used as inputs to the model during training. For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus. We did not find scheduled sampling \citep{bengio2015scheduled}---having the model use its own transition decisions sometimes at training time---to help.

@@

\begin{table}[t]
\centering
\begin{tabular}{c @{\hspace*{1.5em}}c @{\hspace*{1.5em}} lc}
  \toprule
  $t$ & $S$[$t$] & \multicolumn{1}{c}{$Q_t$} & $a_t$ \\
  \midrule
  0 &   & \underline{\hphantom{0} \hphantom{0}} & \shift \\
  1 &  \word{Spot} & \hphantom{0} \underline{\hphantom{0} 1} & \shift \\
  2 &  \word{sat} & \hphantom{0} \hphantom{0} \underline{1 2} & \shift \\
  3  &   \word{down} & \hphantom{0} \hphantom{0} 1 \underline{2 3} & \reduce\\
  4  &   (\word{sat down}) & \hphantom{0} \hphantom{0} \underline{1 4} & \reduce\\
  5  & (\word{Spot} (\word{sat down})) & \hphantom{0} \underline{\hphantom{0} 5} &        \\
  \bottomrule
\end{tabular}
\caption{The thin-stack algorithm operating on the input sequence $\mathbf x = \text{(\word{Spot}, \word{sat}, \word{down})}$ and the transition sequence shown in the second column. $S$ is shown in the third column and represents the top of the stack at each step $t$. The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce\ operation at the next step.}
\label{tbl:thin-stack}
\end{table}

@@

A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep, any element of which may be involved in later computations. A na\"ive backpropagation implementation would then require storing each of the $O(N)$ stacks for a backward pass, leading to a per-example space requirement of $O(NTD)$ floats. This requirement is prohibitively large for significant batch sizes or sentence lengths $N$. Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the top of the stack.

@@

\begin{figure}
\centering
\resizebox{1\linewidth}{!}{
\begin{tikzpicture}
\begin{axis}[ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
             xticklabels={0,32,64,128,256,512,1024,2048}, xmode=log, log basis x=2,
             legend style={at={(0.55, 0.5)},anchor=north west}]
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.tsv};
  \addlegendentry{Thin-stack GPU}
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.tsv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.tsv};
  \addlegendentry{RNN}
\end{axis}
\end{tikzpicture}
}
\caption{Feedforward speed comparison.}
\label{fig:speed}
\end{figure}

Each model is restricted to run on sentences of 30 tokens or fewer. We fix the model dimension $D$ and the word embedding dimension at 300. We run the CPU performance test on a 2.20 GHz 16-core Intel Xeon E5-2660 processor with hyperthreading enabled. We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.

@@

SNLI is a corpus of 570k human-labeled pairs of scene descriptions like this one. We use the standard train--test split and ignore unlabeled examples, which leaves about 549k examples for training, 9,842 for development, and 9,824 for testing. SNLI labels are roughly balanced, with the most frequent label, {\it entailment}, making up 34.2\% of the test set.

Although NLI is framed as a simple three-way classification task, it is nonetheless an effective way of evaluating the ability of a model to extract broadly informative representations of sentence meaning. In order for a model to perform reliably well on NLI, it must be able to represent and reason with the core phenomena of natural language semantics, including quantification, coreference, scope, and several types of ambiguity.

@@

\begin{table*}[t]\small
\begin{center}
\begin{tabular}{lrlrrrr}
\toprule
Param.     & Range & Strategy        & RNN       & SP.-PI-NT   & SP.-PI  & SP. \\
\midrule 
Initial LR & $2\times 10^{-4}$--$2\times 10^{-2}$ & \textsc{log} & $5\times 10^{-3}$  & $3\times 10^{-4}$ & $7\times 10^{-3}$  & $2\times 10^{-3}$\\
L2 regularization $\lambda$ & $8\times 10^{-7}$--$3\times 10^{-5}$   & \textsc{log} & $4\times 10^{-6}$  & $3\times 10^{-6}$   & $2\times 10^{-5}$  & $3\times 10^{-5}$\\
Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
Embedding transformation dropout & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
Classifier MLP dropout & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
\bottomrule
\end{tabular}
\end{center}
\caption{
\label{tab:hyperparams}
Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log-uniform. Dropout parameters are expressed as keep rates rather than drop rates.
}
\end{table*}

@@

Table~\ref{tab:results} shows our results on SNLI. For the full SPINN, we also report a measure of agreement between this model's parses and the parses included with SNLI, calculated as classification accuracy over transitions averaged across timesteps.

@@

We suspect that the hybrid nature of the full SPINN model is also responsible for its ability to perform better than an RNN baseline even when its internal parser is relatively ineffective at producing correct full-sentence parses. It may act somewhat like the tree-based CNN, only with access to larger trees: using tree structure to build up local phrase meanings, and then using the tracking LSTM, at least in part, to combine those meanings.

@@

For a more ambitious goal, we expect that it should be possible to implement a variant of SPINN on top of a modified stack data structure with differentiable \textsc{push} and \textsc{pop} operations \citep[as in][]{grefenstette2015learning,joulin2015inferring}. This would make it possible for the model to learn to parse using guidance from the semantic representation objective, which currently is blocked from influencing the key parsing parameters by our use of hard \shift/\reduce\ decisions. This change would allow the model to learn to produce parses that are, in aggregate, better suited to supporting semantic interpretation than those supplied in the training data.

@@

We acknowledge financial support from a Google Faculty Research Award, the Stanford Data Science Initiative, and the National Science Foundation under grant nos.~BCS 1456077 and IIS 1514268. Some of the Tesla K40s used for this research were donated by the NVIDIA Corporation. We also thank Kelvin Guu, Noah Goodman, and many others in the Stanford NLP group for helpful comments.
